Coherent optical association of a single molecule -- Supplemental material

### 2020-11-12 sm.tex
Update on Overleaf.

\documentclass[aps,secnumarabic,amsmath,amssymb,superscriptaddress]{revtex4}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{color}
\usepackage{graphics}
\usepackage[pdftex]{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[colorlinks=true]{hyperref}
\usepackage{listings}
\usepackage{braket}

@@

\begin{document}
\title{Coherent optical association of a single molecule -- Supplemental material}
\author{Yichao~Yu}
\email{yichaoyu@g.harvard.edu}
\gradstudent
\author{Kenneth~Wang}
\gradstudent
\author{Jonathan~D.~Hood}
\affiliation{Department of Chemistry, Purdue University, West Lafayette, Indianna, 47906, USA}
\author{Lewis~R.~B.~Picard}
\gradstudent
\author{Jessie~T.~Zhang}
\gradstudent
\author{William~B.~Cairncross}
\harvardccb
\harvardphysics
\cua
\author{Jeremy~M.~Hutson}
\affiliation{Joint Quantum Centre Durham-Newcastle, Department of Chemistry, Durham University, Durham, DH1 3LE, United Kingdom}
\author{Till Rosenband}
\harvardphysics
\author{Kang-Kuen~Ni}
\email{ni@chemistry.harvard.edu}
\harvardccb
\harvardphysics
\cua

@@

\section{Calculating Matrix Elements with Coupled Channel Ground State}
The free atom state (with spherically-symmetric harmonic confinement) and target weakly-bound molecular state are calculated using coupled-channel methods. Due to the cylindrical symmetry of the system, the projection of the total spin $m_F \equiv m_{F_{\text{Na}}} + m_{F_{\text{Cs}}}$ is conserved in the transition. Since our initial spin state is $ \ket{F_{\text{Na}} = 2, m_{F_{\text{Na}}} = 2, F_{\text{Cs}} = 3, m_{F_{\text{Cs}}} = 3}$, we can couple to all channels with total $ m_F = 5 $. These channels, written in the basis $ \ket{S, m_S, I, m_I}$, include $ \ket{0,0,5,5}, \ket{1,0,5,5}, \ket{1,1,5,4}, \ket{1,1,4,4} $, where $S = 0,1$ is the total electron spin and $ I = 2,3,4,5 $ is the total nuclear spin. The coupled-channel wavefunction contains a component in each of these channels . In this work, we assume the excited state wavefunction is a single-channel triplet wavefunction. Thus, only the triplet component of the coupled-channel ground state contributes to $\Omega_a$ and $ \Omega_m $.

The full calculation using the methods in section \ref{sm:sect_2} is shown in figure \ref{f-sm}. The strong contribution from the atomic threshold arises from the large overlap between the weakly-bound molecular state and the atomic continuum. This leads to a $1/\delta_{\text{thresh}}^2$ dependence of the scattering rate and favors deeply bound states as the intermediate state. The Raman Rabi frequency is larger than the scattering rate only for these deeply bound states.

>>>>

\documentclass[aps,secnumarabic,amsmath,amssymb,superscriptaddress]{revtex4}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{color}
\usepackage{graphics}
\usepackage[pdftex]{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[colorlinks=true]{hyperref}
\usepackage{listings}
\usepackage{braket}
\usepackage{xcolor}
\newcommand\jmh[1]{{\color{magenta} #1}}

@@

\begin{document}
\title{Coherent optical association of a single molecule -- Supplemental material}
\author{Yichao~Yu}
\email{yichaoyu@g.harvard.edu}
\gradstudent
\author{Kenneth~Wang}
\gradstudent
\author{Jonathan~D.~Hood}
\affiliation{Department of Chemistry, Purdue University, West Lafayette, Indiana, 47906, USA}
\author{Lewis~R.~B.~Picard}
\gradstudent
\author{Jessie~T.~Zhang}
\gradstudent
\author{William~B.~Cairncross}
\harvardccb
\harvardphysics
\cua
\author{Jeremy~M.~Hutson}
\affiliation{Joint Quantum Centre Durham-Newcastle, Department of Chemistry, Durham University, Durham, DH1 3LE, United Kingdom}
\author{Till Rosenband}
\harvardphysics
\author{Kang-Kuen~Ni}
\email{ni@chemistry.harvard.edu}
\harvardccb
\harvardphysics
\cua

@@

\section{Computational methods for bound states and scattering}

For scattering and near-threshold bound states, we solve the Schr\"odinger equation by
coupled-channel methods \cite{HUTSON1994}, using a basis set in which the electron spins
$s_{\rm Na}$ and $s_{\rm Cs}$ and the nuclear spins $i_{\rm Na}$ and $i_{\rm Cs}$ are separately
coupled,
\begin{equation}
| (s_{\rm Na} s_{\rm Cs} S)
(i_{\rm Na} i_{\rm Cs} I) F M_F\rangle.
\label{eqbassif}
\end{equation}
Here $S$ and $I$ are the total electron and nuclear spins of the molecule or atomic pair. Their
resultant $F$ is the total spin, with projection $M_F$ onto the axis of the magnetic field. In
principle $F$ can be further coupled to the rotational angular momentum $L$, but in the present
work only functions with $L=0$ are included. In the absence of a magnetic field, $F$ and $M_F$ are
both good quantum numbers, but in a magnetic field only $M_F$ is conserved. The calculations in
this paper use basis sets with all possible values of $S$, $I$ and $F$ (for the required value of
$M_F$) that can be constructed with $s_{\rm Na} = s_{\rm Cs} = 1/2$, $i_{\rm Na} = 3/2$ and $i_{\rm
Cs} = 7/2$.

Scattering calculations are carried out using the MOLSCAT package \cite{molscat:2019,
mbf-github:2020} and bound-state calculations using the related BOUND package
\cite{bound+field:2019, mbf-github:2020}. The calculations use singlet and triplet interaction
potentials for Na+Cs obtained by adjusting the potentials of Docenko \emph{et al.}\
\cite{Docenko:2006} at short range, using the same procedure as for KCs \cite{Groebner2017},
to reproduce the experimental binding energy of the least-bound triplet state \cite{Hood2019}
and the position of the s-wave Feshbach resonance near 864~G \cite{Zhang2020}. The atomic
nuclear hyperfine coupling and Zeeman effects are fully included, using the Hamiltonian described
in \cite{Hutson2008}.

The bound-state calculations were initially used to predict the binding energies of states that lie
within about 1~GHz of each atomic threshold of interest. At zero magnetic field, the states fall
into groups from each hyperfine state of the atomic pair, labeled by $(f_{\rm Na},f_{\rm Cs})$. For
example, for the $(f_{\rm Na},f_{\rm Cs}) = (2,3)$ states that are of principal interest here, each
group comprises 5 zero-field levels with $F=1,2,3,4$ and 5; one such group is spread out over a
range of energies between 290 MHz and 770 MHz below the $(f_{\rm Na},f_{\rm Cs}) = (2,3)$
threshold. In a magnetic field, each zero-field level splits into $2F+1$ Zeeman sublevels labeled
by $M_F$. For each sublevel we can calculate the dependence of the energy on magnetic field, in
order to choose a target state for which the Raman transition frequency depends only weakly on
field.

The state that we ultimately chose as the target state for Raman transfer is the least-bound
$(F,M_F)=(5,5)$ sublevel with $(f_{\rm Na},f_{\rm Cs}) = (2,3)$, which was predicted at a binding
energy 763 MHz below the $(f_{\rm Na},f_{\rm Cs}) = (2,3)$ threshold. As described in the text, it
was subsequently located experimentally at a binding energy of 770.57150(9) MHz.

Multichannel bound-state wavefunctions, with components labeled by $S$, $I$ and $F$, are calculated
using the method of Thornley and Hutson \cite{Thornley1994}. For molecular bound states, they are
obtained on a grid of internuclear distances from 2.4~\AA\ to 50~\AA. For states of confined atomic
pairs, a harmonic confining potential is added to the molecular potentials and the grid is extended
to 3000~\AA.

\section{Calculating Matrix Elements with Coupled Channel Ground State}
To calculate Rabi frequencies for transitions to intermediate vibrational levels of the c$^3\Sigma\ (\Omega=1)$ electronic state, we
obtain their wavefunctions from a single-channel calculation using the potential curve of \cite{Grochola2011}. We
then calculate overlap integrals between this function and the triplet ($S=1$) components of the
coupled-channel wavefunctions for either the target molecular state or the confined atomic state.
These calculations assume that the intermediate state has pure triplet character and neglect its
nuclear spin character, which is unknown.

The full calculation of the Raman Rabi frequency and scattering rate using the methods in section \ref{sm:sect_2} is shown in figure \ref{f-sm}. The strong contribution from the atomic threshold arises from the large overlap between the target molecular state and the atomic continuum. This leads to a $1/\delta_{\text{thresh}}^2$ dependence of the scattering rate and favors using deeply bound states as the intermediate state for the Raman transfer. The Raman Rabi frequency is larger than the scattering rate only for the deeply bound states.
